Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-11-24 src/tex/ms.tex
Add comments

@@

\section{Introduction}
\label{sec:Introduction}
% Motivation
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR does not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons gravity suggests that there is gravitational wave (GW) amplitude birefringence, while GR predicts there is no birefringence.
GW amplitude birefringence is a property of space-time which consists of the enhancement of one GW polarization over the other during the propagation of the wave.
the local modification of the amplitude is small, but it grows with the number of cycles the GW is propagated.

% Previous studies
Previous studies have constrained amplitude birefringence by performing Bayesian inference.
In \citet{Okounkova_2022}, they considered the distribution of observed inclination of the GW events in GWTC-2, the second GW transient catalog, and use the distribution to constrain the amplitude birefringence effect with the assumption of frequency independence in GW amplitude birefringence.
In \citet{Yamada_2020} and \citet{Wang_2021}, they both modeled the waveform under birefringence effect with a simplify assumption and performed Bayesian inference on the events in GWTC-1, the first GW transient catalog.

@@

\section{Results}
\label{sec:Results}
\subsection{Result on GWTC-3}
% Stacked kappa posterior plot
\begin{figure}[h]
    \script{kappa_stacked.py}
    \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
    \caption{
        The posterior of $\kappa$ for all 69 events included in this study.
        Each color represent a different event.
        These posterior are used to calculate the population constraint on $\kappa$ in the following.
    }
    \label{fig:kappa_stacked}
\end{figure}

@@

% Corner plot of mean and standard deviation of kappa
\begin{figure}[h]
    \script{corner_Gaussian.py}
    \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
    \caption{
        The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
        Orange line represent the median of $\mu$ and $\sigma$.
    }
    \label{fig:corner_Gaussian}
\end{figure}

@@

\begin{figure*}[ht]
    \script{violin_kappa.py}
    \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
    \caption{
        The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
        Each violin represents a different event.
        The violins are sorted by the quotient of median and standard deviation of the posterior.
        Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
        Orange horizontal line represent $\kappa=0$.
        This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
    }
    \label{fig:violin_kappa}
\end{figure*}

@@

\begin{figure}[h]
    \script{corner_GW150914.py}
    \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
    \caption{
        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
        Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
        Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
        This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
    }
    \label{fig:corner_GW150914}
\end{figure}

@@

\begin{figure}[h]
    \script{corner_GW190521.py}
    \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
    \caption{
        The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
        Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
        2D contour plot shows $39.35\%$ and $90\%$ confidence level.
        This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
    }
    \label{fig:corner_GW190521}
\end{figure}

>>>>

\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}

@@

\section{Introduction}
\label{sec:Introduction}
% Motivation
Since Einstein proposed his theory of general relativity (GR), it was tested in
a wide range of length scales. After a century, we know that GR does not agree
with quantum theories at some length scale. To unify both theories, we need to
study the possibility of different beyond-GR theories. \kw{I think now it kind
of goes into birefringence too fast. Maybe talk about what effect one may expect
to see in GW in general a bit?} Some beyond-GR theories such as Chern-Simons
gravity suggests that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence. GW amplitude birefringence is a
property of space-time which consists of the enhancement of one GW polarization
over the other during the propagation of the wave. the local modification of the
amplitude is small, but it grows with the number of cycles the GW is propagated.

% Previous studies
Previous studies have constrained amplitude birefringence by performing Bayesian
inference\kw{This sentence looks odd to me}. In \citet{Okounkova_2022}, they
considered the distribution of observed inclination of the GW events in GWTC-2,
the second GW transient catalog, and use the distribution to constrain the
amplitude birefringence effect with the assumption of frequency independence in
GW amplitude birefringence. In \citet{Yamada_2020} and \citet{Wang_2021}, they
both modeled the waveform under birefringence effect with a simplify assumption
\kw{What is the simplify assumption} and performed Bayesian inference on the
events in GWTC-1, the first GW transient catalog.

@@

\section{Results}
\label{sec:Results}
\subsection{Result on GWTC-3}
% Stacked kappa posterior plot
% \begin{figure}[h]
%     \script{kappa_stacked.py}
%     \includegraphics[width=\columnwidth]{figures/kappa_stacked.pdf}
%     \caption{
%         The posterior of $\kappa$ for all 69 events included in this study.
%         Each color represent a different event.
%         These posterior are used to calculate the population constraint on $\kappa$ in the following.
%     }
%     \label{fig:kappa_stacked}
% \end{figure}

@@

% Corner plot of mean and standard deviation of kappa
% \begin{figure}[h]
%     \script{corner_Gaussian.py}
%     \includegraphics[width=\columnwidth]{figures/corner_Gaussian.pdf}
%     \caption{
%         The posterior of $\mu$ and $\sigma$ of the population distribution of $\kappa$.
%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
%         Orange line represent the median of $\mu$ and $\sigma$.
%     }
%     \label{fig:corner_Gaussian}
% \end{figure}

@@

% \begin{figure*}[ht]
%     \script{violin_kappa.py}
%     \includegraphics[width=\textwidth]{figures/violin_kappa.pdf}
%     \caption{
%         The violin plot shows the posterior of $\kappa$ for all 69 events included in this study.
%         Each violin represents a different event.
%         The violins are sorted by the quotient of median and standard deviation of the posterior.
%         Blue horizontal line represent the median value of $\mu$, and blue region represent 1 $\sigma$ confidence level.
%         Orange horizontal line represent $\kappa=0$.
%         This plot shows that there are some events with $\kappa$ distribution significantly deviate from $0$.
%     }
%     \label{fig:violin_kappa}
% \end{figure*}

@@

% \begin{figure}[h]
%     \script{corner_GW150914.py}
%     \includegraphics[width=\columnwidth]{figures/corner_GW150914.pdf}
%     \caption{
%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
%         Colors in the plot represent the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3}, the PE done by us with both frequency independent and dependent birefringence respectively.
%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
%         Note that there is no posterior of $\kappa$ for the PE from LVK, as the LVK PE is based on GR, which does not suggest GW amplitude birefringence.
%         This plot shows that frequency independent birefringence create a degeneracy between $\kappa$ and $\iota$, while the frequency dependent birefringence can break it.
%     }
%     \label{fig:corner_GW150914}
% \end{figure}

@@

% \begin{figure}[h]
%     \script{corner_GW190521.py}
%     \includegraphics[width=\columnwidth]{figures/corner_GW190521.pdf}
%     \caption{
%         The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW190521.
%         Colors in the plot are the PE with GR done by LVK without cosmological reweighing \citep{GWTC-2.1, GWTC-3} and the PE done by us with the frequency dependent birefringence.
%         2D contour plot shows $39.35\%$ and $90\%$ confidence level.
%         This plot shows that frequency dependent birefringence cannot break the degeneracy between $\kappa$ and $\iota$.
%     }
%     \label{fig:corner_GW190521}
% \end{figure}
